---
title: "Data Presentation 3"
author: "Jinho Kim"
date: "2022-12-09"
output: 
  html_document:
    code_folding: hide
runtime: shiny
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

suppressPackageStartupMessages({
  suppressWarnings({
    ##load packages
    library(tidyverse)
    library(reshape)
    library(here)
    library(dplyr)
    library(ggplot2)
    library(hrbrthemes)
    library(lubridate)
    library(maditr)
    library(shiny)
    library(DT)
    library(plotly)
  })
})
theme_set(theme_classic())

rm(list=ls()) #clear data environment

##load data
data <- read.csv(here("D:/Personal/Marquette/2022 Fall/SPRT 6190/data presentation 3/Data/DataDP3.csv"))
```

## CMJ Metrics Definitions

Concentric Impulse: Is the amount of force applied during the upward movement i.e. jump.

Concentric Impulse (100ms): Is the amount of force applied during the first 100ms of the jump.

Concentric Mean Power / Body Weight: Is the amount of work done during the jump divided by body weight. 

Eccentric Mean Power / Body Weight: The amount of work done during the dip phase of the jump divided by body weight.

RSI - Modified: Reactive strength index modified is the ratio between jump height and time to take-off. 

Vertical Velocity @ take-off: The velocity in the vertical direction at the moment of take-off. 

## Metric Thoughts

Personally there isn't enough information to determine whether some metrics are useful or not. EDA indicates that the sport in question maybe soccer, however one of the positions is center Field which isn't a position generally recognized in soccer. If the sport is indeed soccer, some of the metrics may not be as useful for every position but as there isn't a clear indication of what the data is for, it wouldn't make sense to take out data. 


```{r}
## Clean Data

data <-  data %>%
  drop_na(age)

data$date <- mdy(data$date)

data1 <- aggregate(data$value, 
                  FUN = median, 
                  by = list(athletes = data$athletes,
                            date = data$date,
                            metric = data$metric,
                            assessment = data$assessment))

data2 <- left_join(data1, data, by = c("athletes" = "athletes", "date" = "date", "assessment" = "assessment", "metric" = "metric"))
data2 <- data2[,-8]

final <- distinct(data2)
```

```{r, include = FALSE, warning = FALSE}

## Data Management

## Group by assessment
anthro <- final[final$assessment == "Anthropometry",]
speed <- final[final$assessment == "Max Running Speed",]
IFT <- final[final$assessment == "30-15 IFT",]
ham_strength <- final[final$assessment == "Nordic Hamstring Strength",]
max_strength <- final[final$assessment == "Max Strength (TrapBar Deadlift)",]
CMJ <- final[final$assessment == "CMJ (Force Plates)",]

## Wide datasets
wide_speed <- dcast(speed, athletes + date + laterality + position ~ metric, value.var="x")
wide_speed <- wide_speed %>%
  drop_na()
wide_CMJ <- dcast(CMJ, athletes + date + laterality + position ~ metric, value.var="x")
wide_IFT <- dcast(IFT, athletes + date + laterality + position ~ metric, value.var="x")
wide_IFT <- wide_IFT %>%
  drop_na()
wide_ham_strength <- dcast(ham_strength, athletes + date + laterality + position ~ metric, value.var="x")
wide_max_strength <- dcast(max_strength, athletes + date + laterality + position ~ metric, value.var="x")

## Normalization function
minMax <- function(x) {
  ((x - min(x)) / (max(x) - min(x)))*100
}

## Apply normalization to wide data sets for quantiles
speed_norm <- as.data.frame(lapply(wide_speed[,c(5:7)], minMax))
speed_norm$athletes <- wide_speed$athletes
speed_norm$date <- wide_speed$date
CMJ_norm <- as.data.frame(lapply(wide_CMJ[,c(5:10)], minMax))
CMJ_norm$athletes <- wide_CMJ$athletes
CMJ_norm$date <- wide_CMJ$date
IFT_norm <- as.data.frame(lapply(wide_IFT[,c(5:6)], minMax))
IFT_norm$athletes <- wide_IFT$athletes
IFT_norm$date <- wide_IFT$date
ham_strength_norm <- as.data.frame(lapply(wide_ham_strength[,c(5:7)], minMax))
ham_strength_norm$athletes <- wide_ham_strength$athletes
ham_strength_norm$date <- wide_ham_strength$date
max_strength_norm <- as.data.frame(lapply(wide_max_strength[,c(5:6)], minMax))
max_strength_norm$athletes <- wide_max_strength$athletes
max_strength_norm$date <- wide_max_strength$date

recent_speed <- speed_norm %>%
  group_by(athletes) %>%
  slice(which.max(as.Date(date, '%Y/%m/%d'))) %>%
  dplyr::rename("0-10 yd" = X0.10y) %>%
  dplyr::rename("10-30 yd" = X10.30y) %>%
  dplyr::rename("0-30 yd" = X0.30y)

recent_CMJ <- CMJ_norm %>%
  group_by(athletes) %>%
  slice(which.max(as.Date(date, '%Y/%m/%d'))) %>%
  dplyr::rename("Concentric Impulse" = Concentric.Impulse) %>%
  dplyr::rename("Concentric Impulse (100ms)" = Concentric.Impulse.100ms) %>%
  dplyr::rename("Relative Concentric Mean Power" = Concentric.Mean.Power...BM) %>%
  dplyr::rename("Relative Eccentric Mean Power" = Eccentric.Mean.Power...BM) %>%
  dplyr::rename("RSI Modified" = RSI.modified) %>%
  dplyr::rename("Vertical Velocity at Takeoff" = Vertical.Velocity.at.Takeoff)

recent_IFT <- IFT_norm %>%
  group_by(athletes) %>%
  slice(which.max(as.Date(date, '%Y/%m/%d'))) %>%
  dplyr::rename("Final Speed" = Final.Speed)

recent_ham_strength <- ham_strength_norm %>%
  group_by(athletes) %>%
  slice(which.max(as.Date(date, '%Y/%m/%d'))) %>%
  dplyr::rename("Lateral Assymetry" = Lateral.Assymetry..) %>%
  dplyr::rename("Left Hamstring Strength" = Left.Hamstring.Strength) %>%
  dplyr::rename("Right Hamstring Strength" = Right.Hamstring.Strength)

recent_max_strength <- max_strength_norm %>%
  group_by(athletes) %>%
  slice(which.max(as.Date(date, '%Y/%m/%d'))) %>%
  dplyr::rename("Estimated 1RM" = Estimated.1RM) %>%
  dplyr::rename("Relative Strength" = Relative.Strength)

speed_long <- melt(recent_speed, id.vars = c("athletes", "date"))
CMJ_long <- melt(recent_CMJ, id.vars = c("athletes", "date"))
IFT_long <- melt(recent_IFT, id.vars = c("athletes", "date"))
ham_strength_long <- melt(recent_ham_strength, id.vars = c("athletes", "date"))
max_strength_long <- melt(recent_max_strength, id.vars = c("athletes", "date"))

recent_final <- rbind(speed_long, CMJ_long, IFT_long, ham_strength_long, max_strength_long)

## drop decimals to 2
recent_final$value <- as.numeric(round(recent_final$value,2), nsmall=2)

recent_final <- within(recent_final, {   
  Grade <- NA
  Grade[value < 25] <- "Poor"
  Grade[value >= 25 & value < 50] <- "Average"
  Grade[value >= 50 & value < 75] <- "Good"
  Grade[value >= 75] <- "Excellent"
  
})

```

## Predict 30 yard dash time
```{r}
## EDA
dashtime <- final %>%
  filter(final$metric == "0-30y")

## Change athlete numbers
a1 <- final %>%
  filter(final$athletes == "athlete 1")

## Some observations from filtering out athletes with 30 yard dash times
## 1. we see that there dash times for 35 out of 39 athletes. 
## 2 30 yard dash times started being collected on 2017-03-08

```

After examining the data, it is inadvisable to predict 30 yard dash time based on the current data. To predict the 30 yard dash time, there needs to be a rigid schedule for testing to obtain optimal results. The current data given is not consistent in scheduling for testing thus making it difficult to draw any correlations. 


Column {.tabset}
-------------------------------------

### Recent Results
```{r}
ui <- fluidPage(
  sidebarPanel(
  selectInput("athletes",
            "Athletes:",
            choices = unique(recent_final$athletes)),
  selectInput("metric",
            "Metrics:",
            choices = unique(recent_final$variable))
  ),
  mainPanel(
   dataTableOutput("table") 
  )
)

server <- function(input, output, session){
  
  dt <- reactive(recent_final[recent_final$athletes == input$athletes & recent_final$variable == input$metric,]) 
  
  output$table <- renderDataTable(
    DT::datatable(dt())
  )
}

shinyApp(ui, server)
```

Side note: Some athletes may not have assessments on certain metrics. Anthropometry metrics were taken out since comparing athletes with quantiles on anthropometry is not optimal. 

### Data Table
```{r}
ui <- fluidPage(
  sidebarPanel(
  selectInput("position",
            "Positions:",
            choices = unique(final$position)),
  selectInput("metric",
            "Metrics:",
            choices = unique(final$metric))
  ),
  mainPanel(
   dataTableOutput("table") 
  )
)

server <- function(input, output, session){
  
  dt <- reactive(final[final$position == input$position & final$metric == input$metric,])
  
  output$table <- renderDataTable(
    DT::datatable(dt())
  )
}

shinyApp(ui, server)
```

Side note: Assessment input was taken out due to not being able choose metrics just from a chosen assessment plus having the metrics only helps with the overall plot reactivity. 

### Graph
```{r}
ui <- fluidPage(
  sidebarPanel(
  selectInput("player",
            "Players:",
            choices = unique(final$athletes)),
  selectInput("metric",
            "Metrics (x):",
            choices = unique(final$metric))
  ),
  mainPanel(
   plotlyOutput('plot')
  )
)

server <- function(input, output, session){
  
  dt <- reactive(final[final$athletes == input$player & final$metric == input$metric,])
  
  output$plot <- renderPlotly(
    ggplot(dt(), aes(date, x))+geom_line()+geom_point()
  )
}

shinyApp(ui, server)
```

Side note: Again assessment input was taken out due to not being able choose metrics just from a chosen assessment plus having the metrics only helps with the overall plot reactivity. 

Given more time, a slider input for manipulating the date would have been more useful. 